Reconstructing Propositional Proofs in Type Theory

### 2017-10-27 paper/sections/type-theory.tex
Cosmetic changes.

To reconstruct proofs generated by \Metis, we have formalized in type theory the
classical propositional logic in~\cite{AgdaProp}.  A briefly description of it
is presented in \cite{VanDalen1994}. In this formalization, we assume the
principle of excluded third (henceforth \abbre{PEM}) as the only axiom since
\Metis is a prover for classical logic. Assuming \abbre{PEM}, we can justify
refutation proofs by deriving from it the \emph{reductio ad absurdum} rule
(henceforth \abbre{RAA}). The \abbrev{RAA} rule is the formulation of the
principle of proof by contradiction, that is, a derivation of a
contradiction,~$⊥$, from the hypothesis~$¬ φ$, is a derivation of~$φ$.

\begin{notation}
  The $\abbre{List}$ type is the usual inductive and parametric type
  for lists. We formalise theorems in \CPL by an inductive family
  $\_⊢\_$ indexed by a list of propositions and a proposition, that
  is, $Γ ⊢ φ$ represents the syntactical consequence relation between
  a list of premises, $Γ : \List~\Prop$ and a conclusion $φ : \Prop$.
\end{notation}

We implemented in~\cite{AgdaProp} the syntactical consequence relation in a
similar way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken}, formation and elimination
rules for connectives and the PEM axiom as the valid inference rules
in Fig.~\ref{fig:CPL-inference-rules}.

>>>>

To reconstruct proofs generated by \Metis, we have formalized in type
theory the classical propositional logic in~\cite{AgdaProp}.  A
briefly description of it is presented in \cite{VanDalen1994}. In this
formalization, we have to assume the principle of excluded third
(henceforth \abbre{PEM}) as an axiom since \Metis is a prover for
classical logic. Assuming \abbre{PEM}, we can justify refutation
proofs by deriving from it the \emph{reductio ad absurdum} rule
(henceforth \abbre{RAA}). The \abbrev{RAA} rule is the formulation of
the principle of proof by contradiction, that is, a derivation of a
contradiction,~$⊥$, from the hypothesis~$¬ φ$, is a derivation of~$φ$.

\begin{notation}
  The $\abbre{List}$ type is the usual inductive and parametric type
  for lists.
\end{notation}

We formalise the syntactical consequence relation of \CPL by an
inductive family $\_⊢\_$ with two indexes, a set of propositions (the
premises) and a proposition (the conclusion), that is, $Γ ⊢ φ$
represents that there is derivation with conclusion $φ : \Prop$ from
the set of premises $Γ : \List~\Prop$. We implemented
in~\cite{AgdaProp} the syntactical consequence relation in a similar
way as it was presented in~\cite{Altenkirch2015}. For that reason, we
have included structural rules like \emph{weaken}, formation and
elimination rules for connectives and the PEM axiom as the valid
inference rules in Fig.~\ref{fig:CPL-inference-rules}.
